### 2015-07-08 manuscript/defs.tex

\documentclass{article}
\usepackage{longtable}
\pagestyle{myheadings}
\usepackage{color}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}

\begin{document}

\begin{definition}
  A {\it repeat element family descriptor (refd)} is a
  string over the alphabet $\{A,C,G,T,*\}$, describing the contents
  of any string in the family.
\end{definition}

\begin{definition}
  We say an refd $r$ {\it matches} a Genome $G$ at position $i$ if, for
  all $j$ such that $r_j \neq *$, $r_j = G_{i+j}$.
\end{definition}

Example If $r = AA*TT$, and $G = AACTTGGAAGTT$, then $r$ matches $G$
at positions $i=0$ and $i=7$.  If $G=AAATTT$, then $r$ matches $G$ at
positions $i=0$ and $i=1$.

\begin{definition}
  A spaced seed $s$ {\it matches} an refd $r$ at position $i$
  if, for every $0 \leq j < |s|$, if $r_{i+j} = *$ then $s_j = 0$.
\end{definition}

Example: If $s=11011$ and $r = AAAAA*GGG$, then $s$ matches $r$
at positions $i=0$ and $i=3$, but not at any other $i$.

\begin{definition}
  Given a seed $s$ and refd $r$, let $M_s(r)$ be the set of values $i$
  such that $s$ matches $r$ at $i$.
\end{definition}

Let $s=11011$ and $r = AA*CC*GGGGGG$.  Then $M_s(r) = [0, 3, 6, 7]$.

\begin{definition}
  A spaced seed $s$ is consistent with an refd $r$ if
  for every $i$, $0 \leq i < |r|$, there is some
  $i - |s| \leq j \leq i$ such that $s$ matches $r$ at position $j$.
\end{definition}

In otherwords: for any position $i$ of the refd, $i$ must be able to
match the seed to a position of $r$ such that it then covers position
$i$.

Example: The seed $11011$ is compatible with $r=AA*AA*AA$.  (The seed
matches at positions 0 and 3, and all positions are covered by these
two.)  But it is not consistent with $AAAAA**AAAAA$, and there is no
seed that can match this string at any position that can cover $i=5$
or $i=6$.

\begin{observation}
  Let $L$ be the sortest sequence of the values in $M_s(r)$.  Then $s$
  is consistent with $r$ if and only if $\max_{0 < j < |s|-1}
  L[j+1]-L[j] \leq |s|$.  In the sorted list, every pair of adjacent
  elements must be within $|s|$ of eachother.
\end{observation}

\begin{definition}
  Given a fixed genome $G$, and fixed spaced seed
  $s$, and a fixed value $f$, we define a elementary repeat family as
  a set $S$ of genome coordinates, $|S| \geq f$, such that there
  exists an refd $r$ where:
  \begin{itemize}
  \item $r$ matches the sequence of length $|r|$ starting at each
    element of $S$.  ($S$ is the set of all instances.)
  \item $s$ is consistent with $r$.  ($r$ corresponds to the seed.)
  \item There does not exist a proper substring $r'$ of $r$ such that $r$
    is consistent with $s$ and $M_s(r') - M_s(r) = \emptyset$.  You
    can't have a proper substring of $r$ that describes sequences
    outside of the instances described by $r$.  (Minimality.)
  \item There does not exist a substring $r'$ propertly containing
      $r$ such that the instances defined by $M_s(r')$ contain all the
    instances of $M_s(r)$.  (Maximality.)
  \end{itemize}
\end{definition}

\end{document}