### 2015-07-09 manuscript/defs.tex

\documentclass{article}
\usepackage{longtable}
\pagestyle{myheadings}
\usepackage{color}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}

\begin{document}

\section{Basic definitions}

My first attempt at providing a coherent mathematical definition of a
repeat element family in the context of a spaced seed.

\begin{definition}
An $L$-mer is a string of length $l$.
\end{definition}

\begin{definition}
Given a spaced seed $s$ of length $l$ and an $l$-mer $x$, the {\it seeded
$l$-mer} $s'$ is created by removing all characters of $x$
corresponding to $0$'s in $s$.  Let $\sigma_s(x)$ denote the seeded
$l$-mer for $s$.
\end{definition}
Example: for $s=11011$ and $x = AACGG$, $\sigma_s(x) = AACC$.

\begin{definition}
  A {\it repeat element family descriptor (refd)} is a
  string over the alphabet $\{A,C,G,T,*\}$, describing the contents
  of any string in the family.
\end{definition}

\begin{definition}
  Given refds $r$ and $r'$, we will say $r'$ is a {\it slack
    substring} of $r$ (denoted $r' \prec r$) if there is a
  substring $r''$ of $r$ such that (1) $|r'|=|r''|$, and (2) for all
  $0 \leq i < |r'|$, either $r'_i = r''_i$, $r'_i = *$, or $r''_i=*$.
\end{definition}
(In otherwords, its a substring, with a potentially different *
pattern.)
{\bf Note: } I'm not sure if we should be allowing $r''_i=*$.

\begin{definition}
  We say an refd $r$ {\it matches} a genome $G$ at position $i$ if, for
  all $j$ such that $r_j \neq *$, $r_j = G_{i+j}$.
\end{definition}

Example: If $r = AA*TT$, and $G = AACTTGGAAGTT$, then $r$ matches $G$
at positions $i=0$ and $i=7$.  If $G=AAATTT$, then $r$ matches $G$ at
positions $i=0$ and $i=1$.

\begin{definition}
  A spaced seed $s$ {\it matches} an refd $r$ at position $i$
  if, for every $0 \leq j < |s|$, $s_j = 0$
  whenever $r_{i+j}=*$.
\end{definition}

Example: If $s=11011$ and $r = AAAAA*GGG$, then $s$ matches $r$
at positions $i=0$ and $i=3$, but not at any other $i$.

\begin{observation}
$s$ matches $r$ at $i$ if $\sigma_s(r[i:i+|s|])$ does not contain any
  $*$ symbols.
\end{observation}

$\sigma(r[3:8]) = AAGG$ (hence a match at $i=3$), but $\sigma(r[2:7])
= AA*G$ (hence no match at $i=2$).

\begin{definition}
  Given a seed $s$ and refd $r$, let $M_s(r)$ be the set of values $i$
  such that $s$ matches $r$ at $i$.
\end{definition}

Let $s=11011$ and $r = AA*CC*GGGGGG$.  Then $M_s(r) = [0, 3, 6, 7]$.

\begin{definition}
  A spaced seed $s$ is consistent with an refd $r$ if
  for every $i$, $0 \leq i < |r|$, there is some
  $i - |s| \leq j \leq i$ such that $s$ matches $r$ at position $j$.
\end{definition}

In other words: for any position $i$ of the refd, we must be able to
match the seed to a position of $r$ such that it then covers position
$i$.

Example: The seed $11011$ is compatible with $r=AA*AA*AA$.  (The seed
matches at positions 0 and 3, and all positions are covered by these
two.)  But it is not consistent with $AAAAA**AAAAA$, and there is no
seed that can match this string at any position that can cover $i=5$
or $i=6$.

\begin{observation}
  Let $L$ be the sortest sequence of the values in $M_s(r)$.  Then $s$
  is consistent with $r$ if and only if $\max_{0 < j < |s|-1} L[j+1]-L[j] \leq |s|$.
\end{observation}
That is, in the sorted list, every pair of adjacent elements must be
within $|s|$ of each other.

\begin{definition}
  Given a fixed genome $G$, and fixed spaced seed
  $s$, and a fixed value $f$, we define a elementary repeat family as
  a set $S$ of genome coordinates, $|S| \geq f$, such that there
  exists an refd $r$ where:
  \begin{itemize}
  \item $r$ matches the sequence of length $|r|$ starting at each
    element of $S$.  ($S$ is the set of all instances.)
  \item $s$ is consistent with $r$.  ($r$ corresponds to the seed.)
  \item There does not exist an refd $r'$, $r' \prec r$,  such that $r'$
    is consistent with $s$ and $M_s(r') - M_s(r) \neq \emptyset$.  (You
    cannot have a proper substring of $r$ that describes sequences
    outside of the instances described by $r$ -- minimality.)
  \item There does not exist an refd $r'$, $r \prec r'$, such that $s$ is consistent
    with $r'$, such that the instances defined by $M_s(r')$ contain all the
    instances of $M_s(r)$.  (Maximality.)
  \end{itemize}
\end{definition}
 {\bf Comment:} I'm not sure if the $r \prec r'$ is the right
 relationship.  Maybe just straight substring?  Or perhaps the
 definition of $\prec$ isn't quite right?

\begin{definition}
Given the refd $r$ of an elementary repeat family with set $S$, we say
that $r$ is {\it tight} if, for each $i$ such that $r_i = *$, there
exists two sequences defined by $S$ that have different bases in
position $i$.
\end{definition}
In other words: $r$ is tight if it only uses $*$ symbols where it must
to match everything sequence defined by $S$.

{\bf Comments:} I had it in mind that the algorithm would always
return a tight refd.  But in Carly's defense she gave the example:
s=11011, AAAAATCCCCC, AAGAATCCGCC, where ends up with AA*AA*AA*AA,
which is not tight.  Interestingly, if we have AAAAATTCCCCC and
AAGAATTCCGCC, then we get AA*AATTCC*CC -- to that extra T makes it
tight.  Not sure if this is significant.

\section{From Thesis}

In the following I'm going through definitions / parallels from Nate's
thesis and seeing if I can create an analgy for spaced seeds.

The following are things I wanted to try to prove that may or may not
be useful.  In all of these I'm assuming a fixed genome $G$, a fixed
frequence requriement $f$, and a fixed seed $s$ with length $l$ and
weight $w$.

\begin{lemma} 
For any $l$-mer $x$, the seeded $l$-mer $\sigma_s(x)$ can be a member
of at most $l$ different families.
\end{lemma}
My gut is that this is true, but not necessarily tight.  Maybe its
$w$, or $l-w$?

\begin{definition}
Let $x$ and $y$ be two strings such that $x = a \dot b$, $y = b \dot
c$, and $|b| = i$.  Then $x \circ_i y = a\dot b \dot c$.
\end{definition}

$AAACC \circ_2 CCGGG = AAACCGGG$.  $AAACC \circ_3 CCGGG$ is undefined.

This is a modification of Nate's $x \circ y$ operator.  May need to be
adapted for seeds.

\section{Random Lemmas}

\end{document}