### 2015-09-18 manuscript/JCB/manuscript.tex
Updating manuscript -- copied over text, worked on algorithm format.

\documentclass[11pt]{article}
\usepackage{longtable}
\usepackage{ulem}
\usepackage{authblk}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{algorithmicx}
\usepackage{color}
\def\includegraphic{}
\def\includegraphics{}
\pagestyle{myheadings}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\cyan}[1]{{\color{cyan}#1}}
\newcommand{\magenta}[1]{{\color{magenta}#1}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\usepackage[paper=letterpaper,margin=1in]{geometry}

\title{phRAIDER: Pattern-Hunter based Rapid Ad Initio \\ Identification of Elemenrary Repeats}
\author[1]{Carly E. Schaeffer}
\author[1]{Nathniel D. Figueroa}
\author[2]{Xiaolin Liu}
\author[1,2,3,4]{\\John E. Karro\thanks{Corresponding Author}}
\affil[1]{Department of Computer Science and Software Engineering}
\affil[2]{Cell, Molecular, and Structural Bology}
\affil[3]{Department of Microbiology}
\affil[4]{Department of Statisticcs, Miami University, Oxford, Ohio (USA)}
\renewcommand\Affilfont{\itshape\small}

\newcommand{\comment}[1]{\State \red{\# #1}}

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}

\date{}

\begin{document}
\maketitle
\abstract{This is the abstract.}

\section*{Introduction}

Transposable Elements (TEs) are genomic sequences thathad the capacity
to insert copies of themselves into other genomic locations, resulting
in homologous families of sequences spread across the genome.  Present
in almost every higher order genome (covering as much as 45\% of the
human genome and 90\% of the maize genome
\cite{Venter:2001p92,SanMiguel:1996wa}), TEs have proved an important
source of data in numerous studies of genomic structure (e.g.
\cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97}).
But given their prevalence, it is important for those studying other
aspect the genome to have TEs masked out -- their bases replaced by N
to allow for easy identification and filtering.  Failure to filter can
reek havoc with genomic analysis tools.  For example, unfiltered TEs can trigger
huge numbers of false positives in automated gene finding tool
\cite{Jiang:2013jt}, as well as inflate tool runtime.

The best tools for repeat identification are RepeatMasker and nHMMER
\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ 
library-based search strategies using a pre-compiled
description of sequences in the family (e.g. a ancestral sequence for
BLASTing, or a profile HMM).  But, much like we ask how
the snow plow driver gets to work \cite{Pratchett:uw}, we must ask how
these libraries are compiled.  Library-based tools
are useless for the discovery of new families, and this are
challanging to use on newly sequenced genomes.

Within mammalian species we can largely rely on homology relationships
to port libraries across species.  This does not hold
so well in plants: in many cases TE composition of a given plant
organism is species-specific.  For example, a rice-based TE library
will only identify 25\% of the TEs in the maize genome
\cite{Jiang:2013jt}.

To solve this problem we turn to {\it de novo} TE identification
tools, identifying TEs using only the genome sequence information.  A
number of such tools are discussed in the literature.  RECON uses
WU-BLAST and PILER using LASTZ to compute self-alignments
\cite{Bao:2002,Edgar:2005p2365,Lopez:2003td,Harris:2007uf}.  RECON
show good sensitivity but is computationally intensive and infeasible
for use on whole genomes (requiring 60 hours for 18Mb rice genome in a
2013 study), while PILER achieves a good runtime with very low
sensitivity .  ReAS and RepeatScout \cite{Price:2005p1247} are based
on $k$-mer searches, with the earlier showing less sensitivity than
RECON \cite{Li:2005he, Price:2005p1247,Jiang:2013jt}.  RepeatGluer
\cite{Pevzner:2004p3157,Zhi:2006p3199} is based on a variation of
DeBrujin graphs, which allows for a decomposition of TE families into
domains, but is very, very computationally expensive.  In Saha {\it et
  al.} the authors perform an extensive comparison of the tools, and
conclude that RepeatScout is the best tool overall for assembled
genomes, while ReAS the best when dealing with unassembled sequence
fragments \cite{Saha:2008dm}.

\subsection*{Elementary Repeats}
Zheng and Lonardi approached the {\it de novo} identification problem using {\it elementary repeats}
\cite{Zheng:2005bl}.
Similar to the RepeatGluer domains, elementary repeats are
decompositions of TEs into basic building blocks.  Identification of
these building blocks are sufficient for the purpose
of masking, and can be assembled into Transposable Elements for those
interested in TEs themselves.

While it is notoriously difficult to mathematically model transposable
elements \cite{Bao:2002}, elementary repeats are more conducive to a
formal description.  For a given genome, a nucleotide sequence $r$ is
an elementary repeat if: (1) It is of at least length $l$ (the length
requirement); (2) there are at least $f$ copies of $r$ appear (the
frequency requirement) (3) there is no proper substring of $r$ of
length $\leq l$ that appears in the genome independently of $r$ (the
minimality requirement); (4) $r$ is a maximal string w.r.t (1-3) (the
maximality requirement).  Having proposed this definition, Zheng and
Lonardi developed an identification algorithm that had a runtime
quadratic in the query sequence size \cite{Zheng:2005bl}. This was
refined to linear time by He and also by Huo {\it et al.}
\cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
approaches, but these appraoches are limited in their ability to handle sequence
variation.  As we are looking a genome size inputs, and TEs
transposable elements naturally suffer from copy mistakes and
accumulate instance-specific base substitutions over time, this is a
significant limitation.

\subsection*{RAIDER}
\label{RAIDERSec}
It was with the objective of creating a linear time identification
algorithm that could handle variation through use of
PatternHunter-like spaced seeds that we developed the prototype RIADER
\cite{Li:2004wl}.  A rough implementation was first presented in
Figueroa {\it et al.}, with more details in the Figueroa masters
thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along
an alternate, but equivalent definition of elementary repeats based on
$l$-mers (sequences of length $l$).  Specifically, it was observed
that the minimality condition could be rewritten as: There is no
$l$-mer contained within elementary repeat $r$ that appears in the
genome more times than $r$.  From there we make four core observations
that form the basis for the RAIDER algorithm: (1) An $l$-mer cannot
belong to two different elementary repeats; (2) Any $l$-mer in the
genome that occurs $f$ or more times is either an elementary repeat or
belongs to one; (3) any two $l$-mers belonging to an elementary repeat
must appear the same number of times in the genome; (4) If two
sequences in the genome that are {\it maximally identical} (that is,
cannot be extended in either direction and still be identical), these
sequences cannot belong to a larger elementary repeat.  (By ``belong''
we mean ``is a substring of'' -- a definition we will be generalizing
shortly.)  For discussion and proof, see the Figueroa Thesis
\cite{Figueroa:2013cz}.

Based on these observations, we discover we can find all elementary
repeats in a single scan of the genome.  Specifically, as we scan from
left to right, we track $l$-mer occurrences and identify multiple
copies of the same $l$-mer.  When we find the same sequence of $l$-mers
occurring multiple times in a row, we can mark it as a tentative family,
then break it down later if we discover violations of the minimality
condition.  The algorithm is summarized in Figueroa {\it et al.}, and
discussed in detail in the Figueroa Thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.

Results of the preliminary implementation were promising. On human
chromosome 22 we saw a $12\times$ speedup over RepeatScout to RAIDER
(2344 seconds to 192 seconds), while coverage of the RepBase
\cite{Jurka:2005bl} ancestral sequence improved (77\% to 84\%), while
on mouse chromosome 19 we saw the same speedup with a significant drop
in coverage (53\% to 30\%).  On the full human genome RAIDER ran in
$6.3$ hours, while RepeatScout was unable to complete it run.  For
details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.

\subsection*{Spaced Seeds}

PatternHunter, a very successful augmentation to BLAST
\cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
  spaced seeds}: improving the sensitivity of string matching based
algorthms by allowing wild-cards in the match.  That is, instead of
requiring two strings matching in 12 consecutive characters, we might
instead require two six-character exact matches seperated by one base
which may or may not match (represented by the {\it seed pattern}
11111101111111), or perhaps three consecutive four-character exact
matches sperated by two bases each (1111001111001111). It has been
demonstrated that certain seed patterns can indunce significant
imporvements in BLAST sensitivity with out time penalty, though what
makes a good pattern is not well understood.

RAIDER was designed with the intent of employing the spaced seed
strategy, but this was only implemented heuristically for the Figuera
{\it et al.} paper \cite{Figueroa:2013cz} -- serving primarily as a
proof-of-concept.  Since its release we have develoed a formal model
of elementary repeats that incorporated spaced seeds, and from that
developed phRAIDER (PatternHunter-based RAIDER).  phRAIDER is a fast
tool for the identification and making of transposible elements in
both assembled and unassembled genomes aht outpreforms RptScout and
other establihsed tools. Code is free available under the Gnu GPL
lisence (v. 3) and  may be obtained [NEED WEB ADDRESS].

In the following we will present a new theoretical model for {\it
  seeded elementary repeats}, followed by an algorithm for identifying
them.  In the results section we will show the result of using that
algorithm for masking transp

In our Methods section
we present the new model that allows us to extend RAIDER to correctly
use spaced seeds, and briefly outline the algorithm (with more details
provided in the supplementary materials), with a analysis of phRAIDER
performance in our Results.

\section*{Model}

Our goal is to redefine the concept of transposbile elements to
accomodate a spaced seed strategy.  In the next section we will describe our
identification algorithm, and then quantified its success in 
masking transposable elements. But we will start here with a brief
outline of our theoretical model (with a more detailed description in
the appendix).

We first need to redfine out terminimology regarding elementary
repeats.  Under the Z\&L definition, all instances of one elementary
repeat have the exact same sequence, and hence can be descrbied by a
single string.  As we are allowing for variation, we will describe out
instance set with a {\it sequence descriptor} consisting of bases
letters and the wild-card character * (e.g. AAC*GG would describe a
set of sequences with starting with an AAC, ending with a GG, and
having any character in between).  Given a binary string $s$
representing a spaced seed, we say a sequence descriptor $r$ is
{\it consistent} with $s$ if we can align $s$ to a substring of $r$ such
that every $*$ in $r$ aligns with a $0$ character in $s$. (Hence
$ACG**T*A$ is consistent with the seed $11001$, but not $11011$.)

Given a sequence descriptor $r$, we can {\it decompose} $r$ with
respect to $s$ by taking every length $|s|$ substring of $r$ that is consistant
with $s$, replacing all letters of $r$ that match to a 0 in $s$ with a
$*$, and creating a set from the results.  (Hence for $s = 11011$ and
$r=AA*CCGTT$, the decomposition would be $\{AA*CC, CC*TT\}$.)  We say
$s$ {\it covers} $r$ is every base in $r$ is contained in at least on
string in the decompositio.  (Hence the previous $s$ does not cover $AAA*CC$, as
the first base in not in any of the strings of the decomposition.)

We can now modif the previous definition of elementary repeats as
follows:
\begin{definition} Given a genomic sequence $G$, an integer $f$, and a
  spaced seed $s$, a sequence descriptor $r$ describes an elementary
  repeat if it meets the four (moified) requirements of an elementary repeat:
\begin{itemize}
\item \underline{Structure requriement}: $s$ covers $r$.
\item \underline{Frequency requirement}: There are at least $f$
  substrings of $G$ that match $r$.
\item \underline{Minimality requirement:} For every string $t$ in the
  decomposition of $r$ w.r.t $s$, the number of occurences of $t$ in
  $G$ is equal to the number of occurneces of $r$ in the genome.
\item \underline{Maximality requirement:} There is no sequence
  descriptor $r'$ of $r$ that contains $r$ as a proper substring and
  satisfies conditions 1-1.
\end{itemize}
\end{definition}

\begin{theorem} When the seed $s$ has no 0 characters, this
    definition of elementary repeats is equivilent to the Z\&L
    definition.
\end{theorem}

\subsection*{phRAIDER Algorithm}

\begin{algorithm}
\caption{This is the algorthm caption.}
\begin{algorithmic}
  \Function{sameOffset}{$l$-mer $v_1$, $l$-mer $v_2$}
    \comment{Is the distance between the first two occurences of $v$ equal to the}
    \comment{the distance betwen the most recent two occurences?}
    \State \Return $H[v_2][0] - H[v_1][0]  == H[v_2][-1] - H[v_1][-1]$
  \EndFunction
  \State
  \Function{phRAIDER}{Genome $G$, Seed $s$, MinFrequence $f$}
    \comment{We are going to maintain a $Q$ of up to $|s|$ families, representing the families}
    \comment{of the last $|L|$ $l$-mers seen, sorted by the position of those last $l$-mers.}
    \State Queue$<$Family$>$ $Q$;
    \State
    \comment{Iterate over all $l$-mer positions in $G$}
    \For {$i \leftarrow 1 \to |G|-|s|$}

       \comment{Remove any families whose last $l$-mer did not occur in the last $i$ bases.}
       \While{$i - $location($Q$, finish, last, end)$ > |s|$}
         \State $Q.$dequeue()
       \EndWhile
       \State
       \comment{Get the $l$-mer occuring at $G_i$ and remove letters corresponding to $0$s in $s$.}
       \State $v \leftarrow \mbox{seeded}(G[i],x)$ 
       \State
       \comment{Add to list of $v$'s positions.}
       \State $H[v].push(i)$
       \State
       \If {$|H|(v)| == 2$}
          \comment{$v$ either is the beginning of a family, or should be combined with a}
          \comment{previous $l$-mer to form a family.}
          \State $F = \mbox{arg\_find}_{F \in Q}\{\mbox{sameOffset}(F.\mbox{start}(), v)\}$
          \If {$F$}
             \State $Q.$requeue($F$)
          \EndIf
          \State last($Q$).addLmer($v$)   
       \ElsIf {$|H[v]| > 2$}
          \comment{Is $v$ the proper continuation of a family, or does it need to be split off?}
          \State $F \leftarrow v.$family()
          \If {$H[F.\mbox{lastSeen}(i)][-1] < H[Q.\mbox{front}().\mbox{lastSeen}()]$}
              \comment{$v$ belongs to a family whose last instance is too far back.}
              \If {$v \neq F.$first$()$}
                  \comment{$v$ is not the first $l$-mer of its family}
                  \State $F' \leftarrow F.$split($F.v$)
              \Else
                  \State $F' \leftarrow F.$split($F.$lastSeen())
              \EndIf
              \State $Q.$enqueue($F'$)
          \Else
              \If {$v == F.$first()}
                 \comment{$v$ is the first $l$-mer of $F$, so $F$ is moved to the back of $Q$}
                 \State $Q$.requeue($F$)
              \ElsIf {sameOffSet($F.$first(), $v$)}
                 \State $F.$setLastSeen($v$)
              \EndIf
           \EndIf                             
       \EndIf
       \EndFor
       \State
       \comment{Final break up of families as needed; filter families with frequency $\leq f$.}
       \State tieLooseEnds()
  \EndFunction
\end{algorithmic}
\end{algorithm}

\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{bmc}      % Bibliography file (usually '*.bib' )

\end{document}